Topic modeling bla bla

### 2017-06-06 thesis_report/draft_thesis.tex
Add references

Like any discrete distribution, a multinomial has to sum to one, where the sum
is over all possible data points. Here, a data point is a document containing
$n$ words.

@@

Latent Semantic Indexing was one of the earliest methods for finding
relationships between documents and the words that occur in them. In LSI a
term-document matrix is created by analyzing the corpus, where the rows
correspond to words and columns to documents. Each element in this sparse
matrix describes the number of times a word occurs in a document but this term
count can also be weighted with for instance TF-IDF. Letting each word
represent a dimension in a very high dimensional space, a document can be seen
as a vector with components corresponding to its weighted term counts.

@@

Probabilistic Latent Semantic Indexing evolved from LSI and uses the same
concept of finding a lower rank approximation of the term-document occurrence
matrix. The difference is that instead of being based on linear algebra, PLSI
is based on a mixture decomposition using a latent class model. PLSI associates
an unobserved class variable $z$ with each document-word observation pair $(w,
w)$. This $z$ can be seen as a topic, since it is a probability distribution
over words. As a generative model, PLSI can be defined in the following way for
a corpus $D$:

@@

LDA is a topic model that was first presented as a graphical model for topic
discovery by David Blei, Andrew Ng, and Michael I. Jordan in 2003. The basic
idea is that documents exhibit multiple topics. LDA is a generative statistical
model that allows sets of observations to be explained by unobserved groups
that explain why some parts of the data are similar. For example, if
observations are words collected into documents, it posits that each document
is a mixture of a small number of topics and that each word's occurrence is
attributable to one of the document's topics. This is similar to PLSI, except
that in LDA the topic distribution is assumed to have a Dirichlet prior (see
2.3.3.1). In practice, this results in more reasonable mixtures of topics in a
document. The reduced document description result in PLSI is a vector where
each element describes the mixing proportion for a topic. A limitation with
PLSI is that there is no generative model for these proportions, making it
difficult to handle unseen documents. The Latent Dirichlet Allocation model
tries to solve this limitation by setting a Dirichlet prior on the topic
distribution.

@@

The basic idea in LDA is that we define a topic to be a Dirichlet distribution
over a fixed vocabulary. Technically, the model assumes that the topics are
generated first, before the documents. Now for each document in the collection,
we generate the words in a two-stage process:

@@

The Expectation Maximization, or EM, algorithm is an approach to deriving an
algorithm for approximately obtaining maximum likelihood of variables in a
statistical model. Normally one obtains a maximum likelihood solution by taking
partial derivatives of the likelihood function with respect to all variables,
setting these derivatives to zero and solving the equations simultaneously.

@@

Learning the various distributions (the set of topics, their associated word
probabilities, the topic of each word, and the particular topic mixture of each
document) is a problem of Bayesian inference. In the original paper it's being
used a variational Bayes approximation of the posterior distribution;
~\cite{blei2003latent} alternative inference techniques use Gibbs sampling and
expectation propagation. The goal of variational inference is to approximate an
intractable probability distribution $p$, with a tractable one $q$, in a way
that makes them as close as possible given a similarity measure. In this thesis,
variational inference is used to approximate the posterior distribution of
latent variables. A new distribution over the hidden variables $q(z, \beta)$
(called the variational distribution) is  introduced. This new distribution has
properties such that it can be efficiently computed.

>>>>

Like any discrete distribution, a multinomial has to sum to one, where the sum
is over all possible data points. Here, a data point is a document containing
$n$ words ~\cite{Huang_maximumlikelihood}.

@@

Latent Semantic Indexing was one of the earliest methods for finding
relationships between documents and the words that occur in them. In LSI a
term-document matrix is created by analyzing the corpus, where the rows
correspond to words and columns to documents. Each element in this sparse
matrix describes the number of times a word occurs in a document but this term
count can also be weighted with for instance TF-IDF. Letting each word
represent a dimension in a very high dimensional space, a document can be seen
as a vector with components corresponding to its weighted term counts ~\cite{Deerwester90indexingby}.

@@

Probabilistic Latent Semantic Indexing evolved from LSI and uses the same
concept of finding a lower rank approximation of the term-document occurrence
matrix. The difference is that instead of being based on linear algebra, PLSI
is based on a mixture decomposition using a latent class model 
~\cite{Hofmann:1999:PLS:312624.312649}. PLSI associates
an unobserved class variable $z$ with each document-word observation pair $(w,
w)$. This $z$ can be seen as a topic, since it is a probability distribution
over words. As a generative model, PLSI can be defined in the following way for
a corpus $D$:

@@

LDA is a topic model that was first presented as a graphical model for topic
discovery by David Blei, Andrew Ng, and Michael I. Jordan in 2003. The basic
idea is that documents exhibit multiple topics. LDA is a generative statistical
model that allows sets of observations to be explained by unobserved groups
that explain why some parts of the data are similar. For example, if
observations are words collected into documents, it posits that each document
is a mixture of a small number of topics and that each word's occurrence is
attributable to one of the document's topics. This is similar to PLSI, except
that in LDA the topic distribution is assumed to have a Dirichlet prior (see
2.3.3.1). In practice, this results in more reasonable mixtures of topics in a
document. The reduced document description result in PLSI is a vector where
each element describes the mixing proportion for a topic. A limitation with
PLSI is that there is no generative model for these proportions, making it
difficult to handle unseen documents. The Latent Dirichlet Allocation model
tries to solve this limitation by setting a Dirichlet prior on the topic
distribution ~\cite{blei2003latent}.

@@

The basic idea in LDA is that we define a topic to be a Dirichlet distribution
over a fixed vocabulary. Technically, the model assumes that the topics are
generated first, before the documents ~\cite{Blei11introductionto}. Now for each document in the collection,
we generate the words in a two-stage process:

@@

The Expectation Maximization, or EM, algorithm is an approach to deriving an
algorithm for approximately obtaining maximum likelihood of variables in a
statistical model. Normally one obtains a maximum likelihood solution by taking
partial derivatives of the likelihood function with respect to all variables,
setting these derivatives to zero and solving the equations simultaneously 
~\cite{Myung:2003}.

@@

Learning the various distributions (the set of topics, their associated word
probabilities, the topic of each word, and the particular topic mixture of each
document) is a problem of Bayesian inference. In the original paper it's being
used a variational Bayes approximation of the posterior distribution;
~\cite{blei2003latent} alternative inference techniques use Gibbs sampling and
expectation propagation. The goal of variational inference is to approximate an
intractable probability distribution $p$, with a tractable one $q$, in a way
that makes them as close as possible given a similarity measure. In this thesis,
variational inference is used to approximate the posterior distribution of
latent variables. A new distribution over the hidden variables $q(z, \beta)$
(called the variational distribution) is  introduced. This new distribution has
properties such that it can be efficiently computed ~\cite{Fox2011ATO}.
